---
title: "Heat map"
author: "Liliana Cuccu"
date: "March 2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_knit$set(root.dir = "C:/Users/Liliana/OneDrive - Universitat de Barcelona/Projects")
```

This script reports the code used to produce one plot included in the academic paper “Just Reallocated? Robots, Displacement, and Job Quality”, authored by Liliana Cuccu and Vicente Royuela (available [here]( https://www.ifo.de/sites/default/files/events/2022/pillars22-Cuccu.pdf)).

The plot is a heat map representing workers' migration flows between Spanish provinces after a lay-off. Worker-level information is taken from the Muestra continua de vidas laborales (MCVL), an anonymised panel extracted from the Spanish Social Security records, covering the period 2001-2019. The dataset comprises 4% of the reference population, roughly amounting to one million individuals, and provides a detailed set of characteristics for each of their work or unemployment spells, such as start and termination date, cause of contract termination, province of work, economic sector, earnings, contract type, and number of workers employed in the same firm.Provinces in the y-axis represent the province of origin (i.e., the province in which the worker was employed before being laid off) while provinces in the x-axis represent the destination provinces (i.e., the province in which the worker starts a new job). The flows are reported in hundreds of transitions and only include workers estimation sample described in the paper. For each row, column “N” reports the net inflows by province.

```{r libraries, include=TRUE}
# Clean environment
rm(list=ls())

# Libraries
library(gplots)
library(openxlsx)
library(sf)
library(tidyverse)
library(ggspatial)
library(ggplot2)
library(rgdal) 
library(broom) 
library(wesanderson)
library(viridis)
library(RColorBrewer)
library(arules)
library(dplyr)
library(readxl)
```

```{r data_prep, include=TRUE}

# Import data
data <- read_excel("01_Data/Transit_ByProvince_long.xlsx", sheet = "Sheet1" , col_names=TRUE)

# Define order 
lev <-  unique(as.vector(data$orig))
data <- mutate(data, orig2=factor(orig, levels = lev))

lev <-  unique(as.vector(data$dest))
data <- mutate(data, dest2=factor(dest, levels = lev))

# discrete variable
data$p2 <- discretize( data$p, method="fixed", 
                       breaks = c(-1000,-500,-100,-20,0,20,40,60,80,100,500,1000,3000, 
                                  max(data$p[!is.na(data$p)])),
                       labels = c("-1000; -500","-100; -500","-20; -100","-1; -20", "1; 20", 
                                "20; 40", "40; 60","60; 80","80; 100",
                                "100; 500","500; 1000","1000; 3000","3000; 5000"))
```

```{r plot, fig.width=10.8,fig.height=7.6, include=TRUE}
# basic ggplot
par(omi=c(0,0,0,0), mgp=c(0,0,0),mar=c(0,0,0,0)) # margins

pl <- ggplot(
  data,aes(x=dest2,y=orig2,fill=p2)) +
  geom_tile() +
  geom_tile(colour="white",size=0.25)+
  # remove extra space
  scale_y_discrete(expand=c(0,0))+
  # set a base size for all fonts
  theme_grey(base_size=8) +
  # theme options
  theme(
    # set thickness of axis ticks
    axis.ticks=element_line(size=0.4),
    # size axis labels and legend
    axis.text=element_text(size=9.5),
    legend.text=element_text(size=8),
    # size axes titles
    axis.title=element_text(size=10,face="bold", line=8), 
    # remove plot border
    panel.border=element_blank(),
    legend.position="bottom", legend.box = "horizontal", 
    # space between legend and plot
    legend.margin=margin(0,0,0,0),
    # no white margins
    plot.margin=grid::unit(c(0,0,0,0), "mm"),
    # legend style
    legend.key.height= unit(5, 'mm'),
    legend.key.width= unit(17, 'mm'),
    legend.spacing.x = unit(0, 'mm'),
    legend.title.align=0.5,
    legend.title = element_text(size=9, face="bold"),
    # left-align y labels
    axis.text.y = element_text(hjust = 0) ) +
  theme(legend.position = "bottom") +
  
  # axis titles
  labs(y='Origin',x='Destination') +
  # color scale
  geom_polygon(aes(fill=p2), color = "#525252", size = 0.1)+
  scale_fill_manual(values=c('#cb181d','#fb6a4a' ,'#fcae91','#fee5d9',
                             '#ffffd9','#edf8b1','#c7e9b4','#7fcdbb',
                             '#41b6c4','#1d91c0','#225ea8','#253494','#081d58'), na.value="#d9d9d9") +
  guides(fill=guide_legend(title="Flows", label=TRUE, 
         label.position = "bottom", nrow = 1, title.position = "top"))

pl
```